---
title: "Homework-2"
author: "Dheeraj Oruganty"
format: 
    html:
        self-contained: true
---


## Problem-1

Consider the random variable defined by counting the number of failures until the first success, for independent trials with success probability $p$.  Given $p$ between $0$ and $1$, the __R__ commands `myattempts(p)` and `rgeom(1,p)` both simulate this random variable. Find a way of demonstrating that the two commands indeed give the same results, for five different values of p. 
_You may find the table() function useful._

```{r}

set.seed(123)
mytoss = function(p){
  u <- runif(1)
  x <- as.numeric(u < p)#when p is the larger number, there is a higher chance of being this argument true.
  return(x)
}
myattempts = function(p){ counter <- 0
while (mytoss(p) == 0){ counter <- counter + 1 }
return(counter) }

```



Choose five p's of your choice, e.g. $p \in \{.1, .3, .5, .7, .9\}$. Run `myattempts` and `rgeom` each 10000 times. 

a. For the first p, store the fraction of outcomes in the simulation in the columns of a suitable data frame and compare (Hint: use 3 columns to compare `rgeom()`, `myattempts()` and `dgeom()` ) 

```{r}
set.seed(1000)

myattempts_df <- data.frame(replicate(10000, myattempts(0.5)))
myattempts_df<- data.frame(table(myattempts_df[, 1]))

rgeom_df <- data.frame(replicate(10000, rgeom(1,0.5)))
rgeom_df <- data.frame(table(rgeom_df[, 1]))

df_all <- merge(myattempts_df, rgeom_df, by = "Var1", sort = FALSE)

df_all$Var1 <- as.integer(df_all$Var1)-1

x <- as.integer(df_all$Var1)
x

df_dgeom <- data.frame(dgeom(x,0.5))

df_all <- cbind(df_all, df_dgeom)

df_all$Freq.x <- df_all$Freq.x/10000
df_all$Freq.y <- df_all$Freq.y/10000
df_all

```

b. For the second p, Compare distribution by computing statistics such as mean and standard deviation. 

```{r}

meanmyattempts <- mean(replicate(1000,myattempts(0.7)))
sdmyattempts <- sd(replicate(1000,myattempts(0.7)))


meanrgeom <- mean(replicate(1000,rgeom(1,0.7)))
sdrgeom <- sd(replicate(1000,rgeom(1,0.7)))


df1 <- data.frame(
    myattempt <- c(meanmyattempts,sdmyattempts),
    rgeom <- c(meanrgeom,sdrgeom))

rownames(df1) <- c("Mean", "Standard deviation")

colnames(df1) <- c("Myattempts", "Rgeom")

df1
```

c. For the third p, plot both distributions as histograms in the same plot.  

```{r}
library(ggplot2)
par(mfrow = c(1,2))
hist(replicate(1000,myattempts(0.3)), col = 'darkblue', main = 'Plot of myattempts', xlab = 'Failures', ylab = 'Frequency')

hist(replicate(1000,rgeom(1,0.3)), col = 'Red', main = 'Plot of rgeom', xlab = 'Failures', ylab = 'Frequency')



par(mfrow = c(1,1))
```

d. For the fourth p, make side-by-side box plots.  

```{r}

par(mfrow = c(1,2))
boxplot(replicate(1000,myattempts(0.63)), col = 'darkblue', main = 'Plot of myattempts', xlab = 'Failures', ylab = 'Frequency')

boxplot(replicate(1000,rgeom(1,0.63)), col = 'Red', main = 'Plot of rgeom', xlab = 'Failures', ylab = 'Frequency')

```


e. For the fifth p, plot the two empirical distribution functions in the same plot.  

```{r}

# Defining the Data Frame for ECDF

df2 <- data.frame(x = c(replicate(10000,myattempts(0.1)), c(replicate(10000,rgeom(1,0.1)))), group = gl(2, 10000))


ggplot(df2, aes(x=x, col=group)) +
 
# stat_ecdf() function is used to plot ECDF plot
stat_ecdf()


```

## Problem-2 

Consider the following random experiment: draw a uniformly distributed random number $X_1$ from the interval $(0,1)$. Next, draw a uniformly distributed random number $X_2$ from the interval $(0,1+X_1)$, a uniformly distributed random number $X_3$ from the interval $(0,1+X_2)$ and so on until $X_{20}$. 

Use a monte carlo simulation to give an approximate answer to What is the mean value of $X_{20}$? and use a histogram to identify the distribution of $X_{20}$.

$X_1 \sim unif(0,1)$
$X_2 \sim unif(0,1+X_1)$
$X_3 \sim unif(0,1+X_2)$
.
.
.
$X_{20} \sim unif(0,1+X_{19})$

The __R__ command for drawing a uniformly distributed random number from the interval $(0,b)$ is _runif(1,min = 0, max = b)_.


```{r}
nt <- 10000

# Initialize an empty vector to store X20 values
val <- numeric(nt)

# Perform the Monte Carlo simulation
for (i in 1:nt) {
  X <- runif(1, min = 0, max = 1)
  for (j in 2:20) {
    X <- runif(1, min = 0, max = 1 + X)
  }
  val[i] <- X
}

meanval <- mean(val)


hist(val, main = "Distribution of X", xlab = "X20", col = "red")

cat("Estimated mean of X(20):", meanval, "\n")
```

## Problem-3


Suppose that the daily power consumption of a major city, $X$, has a Gamma distribution with shape parameter $r = 4$ and scale parameter $\rho = 2$. Use __R__ to compute the following quantities: \\

a. $Prob(X \le 12)$
```{r}
# exact
sh = 4
sc = 2 

r = 4
rho = 2
prob1 <- pgamma(12, shape = sh, scale = sc)

cat("Prob(X > 12):", prob1, "\n")
```


b. $Prob(X > 5)$  

```{r}

# Probability that X is greater than 5
prob2 <- 1 - pgamma(5, shape = r, scale = rho)
cat("Prob(X > 5):", prob2, "\n")

```

c. $Prob(|X-8|) < 1$  

```{r}
# Probability that |X - 8| < 1 (equivalent to 7 < X < 9)
prob3 <- pgamma(9, shape = r, scale = rho) - pgamma(7, shape = r, scale = rho)
cat("Prob(|X - 8| < 1):", prob3, "\n")
```

d. $z$ such that $Prob(X < z) = .95$  

```{r}
# z such that Prob(X < z) = 0.95
z <- qgamma(0.95, shape = r, scale = rho)
cat("z such that Prob(X < z) = 0.95:", z, "\n")

```

(Hint:$|X-8| < 1$ is equivalent to $7 < X < 9$)


## Problem-4 

Probability theory says that a binomial distribution, $B(n,p)$ is close to that of a normal distribution with mean $np$ and standard deviation $\sqrt{np(1-p)}$, if $np$ and $n(1-p)$ are both sufficiently large, e.g. at least 10.

Check this by plotting both cumulative distribution functions in the same figure, using a staircase plot for the binomial distribution and a line plot for the normal distribution, for three different cases: a case where both $np$ and $n(1-p)$ are large, a case where $np$ is large and $n(1-p) < 10$, and a case where $np < 10$ and $n(1-p) < 10$.


```{r}
# Function to plot binomial and normal CDFs
plot_cdf <- function(n, p) {
  x <- 0:n

  # Compute binomial CDF
  bicdf <- pbinom(x, size = n, prob = p)

  mu <- n * p
  sigma <- sqrt(n * p * (1 - p))

  # Compute normal CDF
  ncdf <- pnorm(x, mean = mu, sd = sigma)

  # Plotting
  plot(x, bicdf, type = "s", main = "Binomial vs Normal CDF",
       xlab = "Number of Successes", ylab = "Cumulative Probability",
       col = "red", lty = 2, lwd = 2)
  lines(x, ncdf, type = "l", col = "gold", lty = 1, lwd = 2)
  legend("right", legend = c("Binomial", "Normal"), col = c("red", "gold"), lty = 1:2, lwd = 2)
}

# Case 1: np and n(1-p) are both large
plot_cdf(n = 50, p = 0.6)

```

In the first case, where np and n(1-p) both are large, the graphs for both of them completely overlap.

```{r}

# Case 2: np is large but n(1-p) < 10
plot_cdf(n = 30, p = 0.8)

```

In the second case, where np is large and n(1-p) < 10, the graphs don’t overlap as much.
```{r}

# Case 3: np < 10 and n(1-p) < 10
plot_cdf(n = 10, p = 0.3)


```


Describe what happens in all three cases. In what sense are the cdf's not close in cases 2 and 3? (Hint: Compare with a cdf of the normal distribution)

A : In the third scenario, when both np and n(1-p) are not significantly large, the graphs exhibit less overlap, primarily in the distribution tails.

Cases 2 and 3 present distinct graph characteristics, featuring variations on a broader scale. Their overlap is notably reduced, and it's worth noting that the binomial distribution's cumulative distribution function (CDF) displays more significant step increments in case 3 compared to case 2.




## Problem-5 

A graphical technique for checking whether a sample has an approximate normal distribution is a "quantile-quantile" plot. The __R__ command is `qqnorm(x)`, where $x$ is the vector of sample values. If the plot is approximately a straight line, then this suggests that the sample comes from a normal distribution. Use the dataset from the `openintro` package to find out which of the the four distributions (three exams and course grade) is the closest to a normal distribution? You can load the dataset with the following commands `library(openintro)` and `data(exam_grades)`. Explore this by making `qqnorm()` and `qqline()` plots of the four distributions How close to straight lines are the plots in each case? How do the plots differ from straight lines? Hint: Make sure to remove the NA values.
```{r}
library(openintro)


data(exam_grades)
cleandata <- na.omit(exam_grades)
head(cleandata)

par(mfrow = c(2,2))


qqnorm(cleandata$exam1)
qqline(cleandata$exam1, col = 2)

qqnorm(cleandata$exam2)
qqline(cleandata$exam2, col = 2)

qqnorm(cleandata$exam3)
qqline(cleandata$exam3, col = 2)

qqnorm(cleandata$course_grade)
qqline(cleandata$course_grade, col = 2)
```

The initial plot exhibits substantial deviations from the straight line. The second plot adheres to the line for the most part but diverges towards the end. In contrast, the third plot deviates from the line and exhibits discrepancies towards the tail. The fourth plot demonstrates better alignment with the reference line compared to the preceding three plots.


## Problem-6 

If $X$ has a continuous distribution with cumulative distribution function $F$, then the new random variable $U = F(X)$ has a uniform $U(0,1)$ distribution. Verify this with simulations for three different continuous distributions of your choice, by making a random sample of sufficient size, sorting it, plugging it into the cdf $F$, and plotting the result.


```{r}

set.seed(432)

expo <- rexp(100000, rate = 2)

expcdf <- pexp(expo, rate = 2)

hist(expcdf, main = "CDF Transformation", xlab = 'CDF Transformation', ylab = 'Frequency', col = 'green')

```

```{r}
normal <- rnorm(100000)

ncdft <- pnorm(normal)

hist(ncdft, main = "CDF Transformation (Normal)", xlab = "CDF Transformation", ylab = "Frequency", col = 'Gold')

```


```{r}
csample <- rcauchy(10000)

ccdft <- pcauchy(csample)

hist(ccdft, main = "CDF Transformation(Cauchy)", xlab = 'CDF Transformation', ylab = 'Frequency', col = 'red')

```


## Optional additional problem

Suppose for $X = X_1 + X_2$ is the sum of two exponentially distributed random variables with the same parameter $\lambda$. Then $X^\alpha$ is very nearly normally distributed for a suitable choice of $\alpha$. Determine an approximate value for $\alpha$ (within 0.05), using a simulation and `qqnorm()` plots for each of your choices of $\alpha$. 

