---
title: "Lab-4: Assignment"
author: "Dr. Purna Gamage"
---

## Problem 1:

Use simulation to answer the following questions.

Given independent uniform random variables $X_1 \sim U(a, b)$ and $X_2 \sim U(a,b)$. (You can use parameters of your choice)

a. Is the distribution of $X_1 +X_2$ uniform? Please Justify your answer.

```{r}
x.4 <- runif(1000)
hist(x.4)

plot.ecdf(x.4)
abline(a=0,b=1, col = 2)

```


```{r}
x.5 <- runif(1000)
x.6 <- runif(1000)
x.7 = x.5 + x.6
hist(x.7)
```


- A: The distribution of $X_1 +X_2$ uniform is not Uniform. When we plot the histogram, we can clearly see that the frequency of the distribution is higher in the middle and decreases gradually.


b. Is the distribution of $X_1.X_2$ uniform? Please Justify your answer.

```{r}
x.5 <- runif(1000)
x.6 <- runif(1000)
x.7 = x.5 * x.6
hist(x.7)

```

- A: The distribution of $X_1.X_2$ is not Uniform. when we plot the histogram, we can clearly see that the frequency of the distribution is strictly decreasing and not uniform.

c. Is the distribution of $max(X_1, X_2)$ or $min(X_1, X_2)$ uniform? Please Justify your answer.

```{r}
x.5 <- runif(1000)
x.6 <- runif(1000)
x.7 = max(x.5,x.6)
hist(x.7)

```

```{r}
x.5 <- runif(1000)
x.6 <- runif(1000)
x.8 = min(x.5,x.6)
hist(x.8)

```

- A: The distribution of $max(X_1, X_2)$ and $min(X_1, X_2)$ is uniform. When we plot the graph, we can see it is uniform.


d. Is the distribution of $1-X_1$ uniform? Please Justify your answer.

```{r}
x.5 <- runif(1000)
x.6 <- runif(1000)
x.8 = 1-x.5
hist(x.8)

```

- A: The distribution of $1-X_1$ is almost uniform. When we plot the graph, we can see the frequency distribution tends to be uniform for all values of x.8

## Problem 2:

Validate the following theorems using simulation. (You can use parameters of your choice)

`optional`: In addition to the simulation, you can also prove these theorems analytically, i.e. with pen and paper, if you want.

a. Linear Transformations 


**Theorem-5.6.4:** If $X$ has the normal distribution with mean $\mu$ and variance $\sigma^2$ and if $Y=a X+b$, 5.6.4 where $a$ and $b$ are given constants and $a \neq 0$, then $Y$ has the normal distribution with mean $a \mu+b$ and variance $a^2 \sigma^2$.

b. Linear Combinations of Normally Distributed Variables

**Theorem-5.6.7:**  If the random variables $X_1, \ldots, X_k$ are independent and if $X_i$ has the normal distribution with mean $\mu_i$ and variance $\sigma_i^2(i=1, \ldots, k)$, then the sum $X_1+\cdots+X_k$ has the normal distribution with mean $\mu_1+\cdots+\mu_k$ and variance $\sigma_1^2+\cdots+\sigma_k^2$.


## Problem 3: 

Use the iris data set to check the normality of Sepal width of setosa using the following normality tests (Repeat the steps  in the example 3 in lab 4, comment  on your results)

a. Anderson Darling Test

b. Kolmogorov-Smirnov Test (Comapre to a standard normal distribution)


## Problem 4: (optional)

`Exponential Distribution`

Given exponentially distributed random variables $X_1, . . . , X_k$ . Think of waiting times for independent random alarm clocks $1,...,k$ to go off.

Which of these are again exponentially distributed? Explore with a simulation.
(Hint: you can use many methods here to compare the distributions, for example; using Kolmogorov-Smirnov Test, plotting cdfs or ecdfs(as we did in lab 3),..etc )

a. Distribution of $min(X_1,\dots,X_k)$? Waiting time for the first alarm to go off.

b. Distribution of $max(X_1, \dots, X_k )$? Waiting time for the last alarm to go off.

c. Distribution of $X_1 + \dots + X_k$ ? Waiting time until Start of the next clock when the previous alarm/s has gone off.