---
title: 'Lab-3'
author: "Dheeraj Oruganty"
format: 
  html: 
    embed-resources: true
---

# Problem 1:

Plot probability density histograms for these discrete distributions (hint: use the d...() function).

a. Binomial Distribution

```{r}
x <- 0:56
p = 0.33
plot(x,dbinom(x,size=100,prob=0.33),type="h", main = "Example of a Binomial distribution")

```




b. Discrete uniform
```{r}
library("extraDistr")

plot(x,dunif(x,min=0,max=50,log=FALSE),type="h", main = "Example of a Discrete Uniform distribution")
```

c. Bernoulli

```{r}
p <- 0.3

x <- 0:1

cdf <- sapply(x, function(k)
pbinom(k, 1, p))

# Create the plot
plot(
x,
cdf,
type = "s",
lwd = 2,
ylim = c(0, 1),
xlab = "Number of successes",
ylab = "Cumulative Probability",
main = "Bernoulli Distribution CDF"
)
```

d. Poisson

```{r}
x <- 0:150
plot(x,dpois(x,lambda=25),type="h", main = "Example of a Poisson distribution")

```

e. Geometric

```{r}
plot(x,dgeom(x,prob=.2),type="h", main = "Example of a Geometric distribution")

```

f. Hyper-geometric

```{r}

plot(x,dhyper(x,m=70,n=30,k=20),type="h", main = "Example of a Hyper-geometric distribution")
```

g. Negative binomial
```{r}
plot(x,dnbinom(x,size=150,prob=.33),type="h", main = "Example of a Negative binomial distribution")
```



# Problem 2.

Suppose there are fifteen multiple choice questions in `DSAN-5100` midterm test. Each question has four possible answers, and only one of them is correct. Find the probability of having four or less correct answers if a student attempts to answer every question at random.

a. Compute the probability of having exactly 4 correct answers by random attempts using _dbinom()_ and _pbinom()_.



b. Find the the probability of having four or less correct answers by random attempts using _dbinom()_.
```{r}

(p4 <- dbinom(4,6,p=0.5) + dbinom(3,6,p=0.5) + dbinom(2,6,p=0.5) + dbinom(1,6,p=0.5) + dbinom(0,6,p=0.5))
```

c. Compute the above probability(part-b) using _pbinom()_.

```{r}
pbinom(4,6,p=0.5)

```


# Problem 3:

`use r to find the probability`

a. Assume an insurance company receives 3 motor vehicle insurance claims per week. What is the probability that they recieve 11 or fewer claims during a month?

$$P(X\le11)?$$

```{r}
a <- ppois(11,3*4)
a
```

b. While you are at the Georgetown library terrace, you notice that airplanes fly at an average rate of 1 every 4 hours. What is the probability that you will see at least one plane in the next hour?


```{r}
1 - ppois(0,0.25)

```

# Problem 4: 

`Try this example with _..nbinom()_`

(This relates to Problem-1 in Lab-1 Assignment)

Mike had the first three successes in trials 6, 8, and 9. He had six failures until he reached three successes.

Do you think Mike has _success probability p = 0.5 or better?_
Can a simulation give an answer?Let's try.

1.  If Mike’s success probability is p = 0.5 What is the probability that he will obtain these 3 successes?. 

```{r}
successes <- 3
p <- 0.5
x <- 6
(prob <- dnbinom(x, size=successes, prob=p))

```

2. Run many simulations (say 10,000) with this success probability to find the same probability $P(X=6)?$ _Hint: Use rnbinom()_

```{r}

simulations <- rnbinom(1000000,size=successes,prob=p)
sum(simulations == 6)/length(simulations)

```

3. If Mike’s success probability were 0.5 or better, he would not need a lot of attempts.
Find the probability that three successes were reached after 9 tosses or later by somebody with success probability 0.5. $P(X\ge6)$

a. Calculate the probabiliy using both _dnbinom()_ and _pnbinom()_.


```{r}
#Mike's success probability
pnbinom1 <- (1 - pnbinom(5,size=successes,prob=p))

dnbinom1 <- (1 - sum(sapply(0:5, function(x) dnbinom(x,size=successes,prob=p))))

pnbinom1
dnbinom1
```

b. Calculate this probability using a Simulation(10,000)

```{r}
#probability using simulations
sim <- rnbinom(1000000,size=successes,prob=p)
sum(sim >= 6)/length(sim)
```

c. Is this probability (part b) the same as you got from "myattempts": Lab 1 Assignment Problem 1 part 3?

No, this is not the same probabability.
